Research Review: \\The Extent of Adaptive Wild Introgression in Crops

### 2018-07-26 paper.tex
updated word count
updated (see slack for details)

Word Count: 3,993

@@

The relationship between maize (\emph{Zea mays} ssp. \emph{mays}) and the teosinte \emph{Zea mays} ssp. \emph{mexicana} (hereafter, \emph{mexicana}) offers a prime example of adaptive wild-to-crop introgression.
Maize was domesticated from \emph{Zea mays} ssp. \emph{parviglumis} (hereafter, \emph{parviglumis})  approximately 9,000 BP in the lowlands of the Balsas River Valley in Mexico \citep{matsuoka2002single}.
From this domestication center, maize spread into the highlands of the Mexican Central Plateau, where it came into sympatry with \emph{mexicana}.
Introgression from \emph{mexicana} to maize has been reported based on both morphological \citep{wilkes1977} and molecular \citep{vanHeerwaarden2011, doebley1987} data.
However, \citet{Hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on ancestry deconvolution methods including HAPMIX (\textbf{Figure 2}).
These introgressed segments overlapped with QTL that had previously been shown to control anthocyanin content and leaf macrohairs \citep{lauter2004}, traits known to be adaptive at high elevation.
In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression had increased height (a proxy for fitness) under highland environmental conditions.
Height differences were not detected under lowland conditions, providing further evidence of local adaptation.

@@

Barley (\emph{Hordeum vulgare} ssp. \emph{vulgare}) was likely domesticated multiple times from wild ssp. \emph{spontaneum} roughly 8,000 to 10,000 BP.
There is clear evidence of one domestication center in the Fertile Crescent \citep{badr2000origin, Morrell2007}, and others have supported additional eastern domestication events, potentially from ssp. \emph{spontaneum} east of the Zagros Mountains \citep{Morrell2007} or from ssp. \emph{spontaneum} var. \emph{agriocrithon} in modern-day Tibet \citep{dai2012tibet}.
However, recent research casts doubt on Tibetan domestication and suggests that var. \emph{agriocrithon} is not a wild relative, but rather a hybrid of domesticated landraces \citep{pourkheirandish2018elucidation}.
Presently, the distribution of wild barley stretches from the eastern Mediterranean to west-central Asia, spanning clines in temperature, precipitation, soil type, and altitude \citep{Morrell2007}.
Cultivated barley is found throughout wild barley's distribution and crop-wild hybrids are fertile and common when these taxa co-occur.

@@

The four crop systems described in detail here represent particularly compelling examples of wild and potentially adaptive introgression.
However, given their similar histories, many additional crops have likely benefited from wild-to-crop gene flow during post-domestication expansion (\textbf{Table 1}).
Across these four cases studies, some generalities can be observed.
Data thus far indicate that wild introgression is often regional in its extent, but that, in certain cases (\emph{e.g.}, \emph{mexicana} haplotypes detected in maize landraces from the Guatemalan or southwestern U.S. highlands), newly introgressed wild haplotypes can be disseminated more broadly.
Additionally, when functional information is evaluated, as in the maize and potato studies, introgression has been found to occur at loci conferring adaptation to novel conditions not found in a crop's center of origin.
The wild gene flow identified in these case studies raises a number of questions regarding both domestication and adaptation of crops.

@@

\subsection*{What candidate genes were targeted by selection during domestication?}
Loci targeted by selection during domestication can be identified through so-called ``bottom-up'' approaches based on population genetic signatures \citep{Ross-Ibarra2007}.
Ideally, candidate loci will be identified by first constructing a demographic model representing the history of the domesticate.
In this approach, polymorphism data from neutral loci are fit to potential models of a crop's demography and then statistical tests of selection are used to identify candidate domestication genes under the most likely model.
Due to the uncertainty associated with any given demography, many studies identify domestication loci using a strict outlier approach in which loci showing, for example, the greatest reduction in nucleotide diversity or the highest allele frequency differentiation in the domesticate relative to the wild progenitor are identified as candidates.
Introgression during crop expansion may influence candidate gene detection using both demographic-modeling and strict-outlier approaches.
For example, \emph{mexicana} introgression into maize described above accounts for approximately 20\% of the genome of maize in the highlands of Mexico \citep{vanHeerwaarden2011}.
Takuno and co-authors \citeyearpar{Takuno2015} have shown that a demographic model incorporating this introgression is a significantly better fit to empirical data than a model lacking introgression.
Failure to account for introgression in maize would therefore compromise domestication candidate detection, particularly if a study contained maize samples from the Mexican highlands where \emph{mexicana} introgression is most prevalent.
Likewise, introgression that increased nucleotide diversity in the domesticate or decreased differentiation at domestication loci would confound a strict outlier approach.
However, previous work, also in maize, has shown that known domestication loci are particularly resistant to wild introgression \citep{Hufford2013}, likely due to ongoing selection favoring the domesticated phenotype.
\vspace{5mm}

@@

\subsection*{Do geographic patterns of introgression inform our understanding of adaptation?}
Conservation of the genomic architecture of introgression across individuals,  between populations, and across landscapes can help illuminate whether introgression is, in fact, adaptive.
For example, if an introgressed chromosomal region is conserved across a broad ecogeographic region, this suggests it may impart adaptation to more widespread environmental or climatic variables (\emph{e.g.}, cool temperatures at high elevation).
On the other hand, if genetic architectures of introgression are conserved across individuals within a population but not across populations in the region, this suggests more local selective pressures (\emph{e.g.}, locally prevalent biotic pressures).
Highly variable introgression across individuals would be more consistent with random gene flow than adaptation.

>>>>

Word Count: 3,966

@@

The relationship between maize (\emph{Zea mays} ssp. \emph{mays}) and the teosinte \emph{Zea mays} ssp. \emph{mexicana} (hereafter, \emph{mexicana}) offers a prime example of adaptive wild-to-crop introgression.
Maize was domesticated from \emph{Zea mays} ssp. \emph{parviglumis} (hereafter, \emph{parviglumis})  approximately 9,000 BP in the lowlands of the Balsas River Valley in Mexico \citep{matsuoka2002single}.
From this domestication center, maize spread into the highlands of the Mexican Central Plateau, where it came into sympatry with \emph{mexicana}.
Introgression from \emph{mexicana} to maize has been reported based on both morphological \citep{wilkes1977} and molecular \citep{vanHeerwaarden2011, doebley1987} data.
\citet{Hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on ancestry deconvolution methods including HAPMIX (\textbf{Figure 2}).
These introgressed segments overlapped with QTL that had previously been shown to control anthocyanin content and leaf macrohairs \citep{lauter2004}, traits known to be adaptive at high elevation.
In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression had increased height (a proxy for fitness) under highland environmental conditions.
Height differences were not detected under lowland conditions, providing further evidence of local adaptation.

@@

Barley (\emph{Hordeum vulgare} ssp. \emph{vulgare}) was likely domesticated multiple times from wild ssp. \emph{spontaneum} roughly 8,000 to 10,000 BP.
There is clear evidence of one domestication center in the Fertile Crescent \citep{badr2000origin, Morrell2007}, and others have supported additional eastern domestication events, potentially from ssp. \emph{spontaneum} east of the Zagros Mountains \citep{Morrell2007} or from ssp. \emph{spontaneum} var. \emph{agriocrithon} in modern-day Tibet \citep{dai2012tibet}.
However, recent research casts doubt on Tibetan domestication and suggests that var. \emph{agriocrithon} is not a wild relative, but rather a hybrid of domesticated landraces \citep{pourkheirandish2018elucidation}.
Presently, the distribution of wild barley stretches from the eastern Mediterranean to west-central Asia, spanning clines in temperature, precipitation, soil type, and altitude \citep{Morrell2007}.
Cultivated barley is found throughout wild barley's distribution \cite{harlan1966distribution} and crop-wild hybrids are fertile and common when these taxa co-occur.

@@

These four crop system case studies represent particularly compelling examples of wild and potentially adaptive introgression.
%The four crop systems described in detail here represent particularly compelling examples of wild and potentially adaptive introgression.
However, given their similar histories, many additional crops have likely benefited from wild-to-crop gene flow during post-domestication expansion (\textbf{Table 1}).
Across these four cases studies, some generalities can be observed.
Data thus far indicate that wild introgression is often regional in its extent, but that, in certain cases (\emph{e.g.}, \emph{mexicana} haplotypes detected in maize landraces from the Guatemalan or southwestern U.S. highlands), newly introgressed wild haplotypes can be disseminated more broadly.
Additionally, when functional information is evaluated, as in the maize and potato studies, introgression has been found to occur at loci conferring adaptation to novel conditions not found in a crop's center of origin.
The wild gene flow identified in these case studies raises a number of questions regarding both domestication and adaptation of crops.

@@

\subsection*{What candidate genes were targeted by selection during domestication?}

Loci targeted by selection during domestication can be identified through so-called ``bottom-up'' approaches based on population genetic signatures \citep{Ross-Ibarra2007}.
Ideally, candidate loci will be identified by first constructing a demographic model representing the history of the domesticate.
In this approach, polymorphism data from neutral loci are fit to potential models of a crop's demography and then statistical tests of selection are used to identify candidate domestication genes under the most likely model.
Due to the uncertainty associated with any given demography, many studies identify domestication loci using a strict outlier approach in which loci showing, for example, the greatest reduction in nucleotide diversity or the highest allele frequency differentiation in the domesticate relative to the wild progenitor are identified as candidates.
Introgression during crop expansion may influence candidate gene detection using both demographic-modeling and strict-outlier approaches.
For example, \emph{mexicana} introgression into maize described above accounts for approximately 20\% of the genome of maize in the highlands of Mexico \citep{vanHeerwaarden2011}.
Takuno and co-authors \citeyearpar{Takuno2015} have shown that a demographic model incorporating this introgression is a significantly better fit to empirical data than a model lacking introgression.
Failure to account for introgression in maize would therefore compromise domestication candidate detection, particularly if a study contained maize samples from the Mexican highlands where \emph{mexicana} introgression is most prevalent.
Likewise, introgression that increased nucleotide diversity in the domesticate or decreased differentiation at domestication loci would confound a strict outlier approach.
However, previous work, also in maize, has shown that known domestication loci are particularly resistant to wild introgression \citep{Hufford2013}, likely due to ongoing selection favoring the domesticated phenotype.
\vspace{5mm}

@@

\subsection*{Do geographic patterns of introgression inform our understanding of adaptation?}
Conservation of the genomic architecture of introgression across individuals,  between populations, and across landscapes can help illuminate whether introgression is adaptive.
For example, if an introgressed chromosomal region is conserved across a broad ecogeographic region, this suggests it may impart adaptation to more widespread environmental or climatic variables (\emph{e.g.}, cool temperatures at high elevation).
On the other hand, if genetic architectures of introgression are conserved across individuals within a population but not across populations in the region, this suggests more local selective pressures (\emph{e.g.}, locally prevalent biotic pressures).
Highly variable introgression across individuals would be more consistent with random gene flow than adaptation.
